
\section*{6.10}

\textbf{Assume two bidders with valuations $v_1$ and $v_2$ for an object. Their bids are in multiples of some unit (that is, discrete). The bidder with higher bid wins the auction and pays the amount that he has bid. If both bid the same amount, one of them gets the object with equal probability $\frac{1}{2}$. In this game, compute a pure strategy Nash equilibrium of the game.}

There are three possible strategies; a bidder $i$ may bid over his own valuation $b_i > v_i$, equal to his valuation $b_i = v_i$ or under his valuation $b_i < v_i$.

Depending on the bid of the other bidders $j$, the utility for $i$ is given as follows
\[
u_i =
\begin{cases}
    v_i - b_i & \text{ if } b_i > b_j \\
    \frac{1}{2}(v_i - b_i) & \text{ if } b_i = b_j \\
    0 & \text{ if } b_i < b_j \\
\end{cases}
\]

There are nine possible outcomes for $i$:
\begin{table}[H]
    \centering
    \begin{tabular}{|c|c|c|c|} \hline
         & $b_i > b_j$ & $b_i = b_j$ & $b_i < b_j$ \\ \hline
        $b_i > v_i$ & $<0$ & $<0$ & $0$ \\ \hline
        $b_i = v_i$ & $0$ & $0$ & $0$ \\ \hline
        $b_i < v_i$ & $>0$ & $>0$ & $0$ \\ \hline
    \end{tabular}
\end{table}

Given that $i$ is an arbitrary bidder, the payoff matrix is the same for all bidders, hence $b_i < v_i$ must be a pure strategy Nash equilibrium.

\section*{6.9}

\textbf{Give examples of two player pure strategy games for the following situations}

\begin{enumerate}
    \item[(a)] \textbf{The game has a unique Nash equilibrium which is not a weakly dominant strategy equilibrium}

    \begin{table}[H]
        \centering
        \begin{tabular}{|c|c|c|} \hline
              & \multicolumn{2}{|c|}{2} \\ \hline
            1 & A & B \\ \hline
            A & 1,0 & 0,1 \\ \hline
            B & 0,1 & 0,0 \\ \hline
        \end{tabular}
    \end{table}
    
    $(A,B)$ is a unique Nash equilibrium.
    
    \item[(b)] \textbf{The game has a unique Nash equilibrium which is a weakly dominant strategy equilibrium but not a strongly dominant strategy equilibrium}

    \begin{table}[H]
        \centering
        \begin{tabular}{|c|c|c|} \hline
              & \multicolumn{2}{|c|}{2} \\ \hline
            1 & A & B \\ \hline
            A & 1,1 & 0,0 \\ \hline
            B & 0,1 & 0,0 \\ \hline
        \end{tabular}
    \end{table}
    
    $(A,A)$ is a unique Nash equilibrium and a weakly dominant strategy equilibrium.
    
    \item[(c)] \textbf{The game has one strongly dominant or one weakly dominant strategy equilibrium and a second one which is only a Nash equilibrium}

    \begin{table}[H]
        \centering
        \begin{tabular}{|c|c|c|} \hline
              & \multicolumn{2}{|c|}{2} \\ \hline
            1 & A & B \\ \hline
            A & 1,1 & 0,1 \\ \hline
            B & 0,1 & 0,0 \\ \hline
        \end{tabular}
    \end{table}
    
    $(A,A)$ is a weakly dominant strategy equilibrium and $(A,B)$ is only a Nash equilibrium.
    
\end{enumerate}

\section*{6.3}

\textbf{Find the pure strategy Nash equilibria, maxmin values, minmax values, maxmin strategies, and minmax strategies of the following game.}

\begin{table}[H]
    \centering
    \begin{tabular}{|c|c|c|} \hline
          & \multicolumn{2}{|c|}{2} \\ \hline
        1 & A & B \\ \hline
        A & 0,1 & 1,1 \\ \hline
        B & 1,1 & 1,0 \\ \hline
    \end{tabular}
\end{table}

\textit{pure strategy Nash equilibrium} $(A,B)$, $(B,A)$

\textit{maxmin values} $\underline{v_1}=1$, $\underline{v_2}=1$

\textit{maxmin strategies} $s_1=\{B\}$, $s_2=\{A\}$

\textit{minmax values} $\overline{v_1}=1$, $\overline{v_2}=1$

\textit{minmax strategies} $s_1=\{A, B\}$, $s_2=\{A,B\}$


\section*{6.1}

\textbf{Show in a strategic form game that any strongly (weakly) (very weakly) dominant strategy equilibrium is also a pure strategy Nash equilibrium.}

Let $s^* = (s_1^*, ..., s_n^*)$ be a dominant strategy equilibrium and $i\in [1,...,n]$ be an arbitrary player. $s_i^*$ is a dominant strategy for $i$, such that
\[ u_i(s_i^*, s_{-i}) \geq u_i(s_i,s_{-i})\ \forall s_{-i}\in S_{-i} \]

Hence
\[ u_i(s_i^*, s_{-i}^*) \geq u_i(s_i, s_{-i}^*) \]
and $s^*$ must also be a Nash equilibrium.